Convergence rate of MAP estimates \\ for the exponential family

### 2021-11-09 aistats.tex
Add ref to prior choice subsection

In this section, we review useful properties of the gamma distribution and associated Bregman divergence in \S\ref{app:gamma}. 
Then we prove theorem~\ref{thm:gammaMLE} in \S\ref{app:gammaMLE}.
Then we prove an extension of theorem~\ref{thm:varianceMLE} in \S\ref{app:multivariateMLE}.
Then we prove a useful lemma about the expectation of the natural parameter of the MAP in \S\ref{app:nat-bound}.
Finally we prove upper bounds for the MAP in \S\ref{app:MAP-bound}.

@@

\begin{theorem}[MAP Bound]
Consider an exponential distribution with sufficient statistics coming from a gamma distribution $\Gamma(\alpha, \beta^*)$ with mean parameter $\mu^* = \frac{\alpha}{\beta^*}$.
The expected symmetrized Bregman~\eqref{eq:symmetrized_bregman} of the MAP  with prior hyper-parameters $(n_0,\mu_0)$ is upper bounded as
\alignn{
    \forall n\geq \inv{\alpha},\quad \quad
	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq\frac{1}{n_0+n} + 
	\frac{\alpha\paren{\frac{\mu_0}{\mu^*} - \inv{\alpha n_0} - 1 }^2 }{\paren{1+\frac{n}{n_0}}\paren{\frac{\mu_0}{\mu^*} +  \frac{n - \inv{\alpha}}{n_0}}}
}
\end{theorem} 
Note that the second term vanishes when $\mu_0 = \mu^* \paren{1 + \inv{\alpha n_0}}$.
Expressions for $n\alpha < 1$ are less elegant. See below \TODO{Damien try this on Matlab?}

\alignn{
 	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq \frac{n_0 + n}{a + n +\inv{\alpha}} \paren{1 + \inv{a \alpha}} -1 + \frac{a - n_0}{n_0+n}
	\quadtext{where}
	 a:=n_0\frac{\mu_0}{\mu^*} \; .
}

@@

\subsection{On the Choice of a Prior}
The optimal $\mu_0$ is larger than $\mu^*$ for small $n_0$ and small $n$.
Indeed, the upper bound~\eqref{eq:MAP_rate} has a bias term that is $0$ when $\frac{mu_0}{\mu^*} = 1 + \inv{n_0}$, e.g. for large values of $n_0$, it is $\mu_0=\mu^*$ is the best prior, but for small $n_0$, one better sets larger values for $\mu_0$. In Figure~\ref{fig:optimal_n0}, we observe this behavior numerically.

>>>>

In this section, we review useful properties of the gamma distribution and associated Bregman divergence in \S\ref{app:gamma}. 
Then we prove theorem~\ref{thm:gammaMLE} in \S\ref{app:gammaMLE}.
Then we prove an extension of theorem~\ref{thm:varianceMLE} in \S\ref{app:multivariateMLE}.
Then we prove a useful lemma about the expectation of the natural parameter of the MAP in \S\ref{app:nat-bound}, in order to prove upper bounds for the MAP in \S\ref{app:MAP-bound}.
Finally we numerically investigate the effect of prior hyper-parameters in \S\ref{app:prior-choice}.

@@

\begin{theorem}[MAP Bound]
Consider an exponential distribution with sufficient statistics coming from a gamma distribution $\Gamma(\alpha, \beta^*)$ with mean parameter $\mu^* = \frac{\alpha}{\beta^*}$.
The expected symmetrized Bregman~\eqref{eq:symmetrized_bregman} of the MAP  with prior hyper-parameters $(n_0,\mu_0)$ is upper bounded as
\alignn{
    \forall n\geq \inv{\alpha},\quad \quad
	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq\frac{1}{n_0+n} + 
	\frac{\alpha\paren{\frac{\mu_0}{\mu^*} - \inv{\alpha n_0} - 1 }^2 }{\paren{1+\frac{n}{n_0}}\paren{\frac{\mu_0}{\mu^*} +  \frac{n - \inv{\alpha}}{n_0}}}
}
\end{theorem} 
Note that the second term vanishes when $\mu_0 = \mu^* \paren{1 + \inv{\alpha n_0}}$.
Expressions for $n\alpha < 1$ are less elegant as visible below
\alignn{
 	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq \frac{n_0 + n}{a + n +\inv{\alpha}} \paren{1 + \inv{a \alpha}} -1 + \frac{a - n_0}{n_0+n}
	\quadtext{where}
	 a:=n_0\frac{\mu_0}{\mu^*} \; .
}

@@

\subsection{On the Choice of a Prior}
\label{app:prior-choice}
The optimal $\mu_0$ is larger than $\mu^*$ for small $n_0$ and small $n$.
Indeed, the upper bound~\eqref{eq:MAP_rate} has a bias term that is $0$ when $\frac{mu_0}{\mu^*} = 1 + \inv{n_0}$, e.g. for large values of $n_0$, it is $\mu_0=\mu^*$ is the best prior, but for small $n_0$, one better sets larger values for $\mu_0$. In Figure~\ref{fig:optimal_n0}, we observe this behavior numerically.
